---
title: "Hands-on Exercise 2: Spatial Weights and Applications"
author: "QIU RUILIU"
date: "22 Nov 2023"
date-modified: "last-modified"
format: html
execute: 
  eval: true
  echo: true
  warning: false
editor: visual
---

# **Ex 2A: Spatial Weights and Applications**

## **Overview**

The main task of this hands-on exercise is to learn how to compute spatial weights using R. We are mainly using:

-   import geospatial data using appropriate function(s) of **sf** package,

-   import csv file using appropriate function of **readr** package,

-   perform relational join using appropriate join function of **dplyr** package,

-   compute spatial weights using appropriate functions of **spdep** package, and

-   calculate spatially lagged variables using appropriate functions of **spdep** package.

## **The Study Area and Data**

Two data sets will be used in this hands-on exercise, they are:

-   Hunan county boundary layer. This is a geospatial data set in ESRI shapefile format.

-   Hunan_2012.csv: This csv file contains selected Hunan's local development indicators in 2012.

### **Getting Started**

Before we get started, we need to ensure that **spdep**, **sf**, **tmap** and **tidyverse** packages of R are currently installed in your R.

```{r}
pacman::p_load(sf, spdep, tmap, tidyverse, knitr)
```

## **Getting the Data Into R Environment**

In this section, we will bring a geospatial data and its associated attribute table into R environment. The geospatial data is in ESRI shapefile format and the attribute table is in csv fomat.

### **Import shapefile into r environment**

The code chunk below uses [*st_read()*](https://r-spatial.github.io/sf/reference/st_read.html) of **sf** package to import Hunan shapefile into R. The imported shapefile will be **simple features** Object of **sf**.

```{r}
hunan <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan")
```

### **Import csv file into r environment**

Next, we will import *Hunan_2012.csv* into R by using *read_csv()* of **readr** package. The output is R dataframe class.

```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```

### **Performing relational join**

The code chunk below will be used to update the attribute table of *hunan*'s SpatialPolygonsDataFrame with the attribute fields of *hunan2012* dataframe. This is performed by using *left_join()* of **dplyr** package.

```{r}
hunan <- left_join(hunan,hunan2012)%>%
  select(1:4, 7, 15)
```

 *select(1:4, 7, 15)*  This function is used to select specific columns from a data frame.  *1:4*  means it selects the first, second, third, and fourth columns. The numbers  *7*  and  *15*  mean it also selects the seventh and fifteenth columns.

## **Visualising Regional Development Indicator**

Now, we are going to prepare a basemap and a choropleth map showing the distribution of GDPPC 2012 by using *qtm()* of **tmap** package.

```{r}
basemap <- tm_shape(hunan) +
  tm_polygons() +
  tm_text("NAME_3", size=0.5)

gdppc <- qtm(hunan, "GDPPC")
tmap_arrange(basemap, gdppc, asp=1, ncol=2)
```
