---
title: "Hands on Exercise 1: Data importing and wrangling"
---

# Overview

In this hands-on exercise, you will learn how to handle geospatial data in R by using **sf** package.

By the end of this hands-on exercise, you should acquire the following competencies:

-   installing and loading [sf](https://r-spatial.github.io/sf/) and [tidyverse](https://www.tidyverse.org/) packages into R environment,

-   importing geospatial data by using appropriate functions of **sf** package,

-   importing aspatial data by using appropriate function of **readr** package,

-   exploring the content of simple feature data frame by using appropriate **Base R** and **sf** functions,

-   assigning or transforming coordinate systems by using using appropriate *sf* functions,

-   converting an aspatial data into a sf data frame by using appropriate function of **sf** package,

-   performing geoprocessing tasks by using appropriate functions of **sf** package,

-   performing data wrangling tasks by using appropriate functions of **dplyr** package and

-   performing Exploratory Data Analysis (EDA) by using appropriate functions from **ggplot2** package.

# **Getting Started**

```{r}
pacman::p_load(sf, tidyverse)
```

# **Importing Geospatial Data**

```{r}
mpsz = st_read(dsn = "D:/KathyChiu77/ISSS624/Hands-on Ex 1/Data/geospatial", 
                  layer = "MP14_SUBZONE_WEB_PL")
cyclingpath = st_read(dsn = "D:/KathyChiu77/ISSS624/Hands-on Ex 1/Data/geospatial", 
                         layer = "CyclingPathGazette")
preschool = st_read("D:/KathyChiu77/ISSS624/Hands-on Ex 1/Data/geospatial/PreSchoolsLocation.kml")
```

# **Checking the Content of A Simple Feature Data Frame**

```{r}
st_geometry(mpsz)
glimpse(mpsz)
head(mpsz, n=5)  
```

# **Plotting the Geospatial Data**

```{r}
plot(mpsz)
plot(st_geometry(mpsz))
plot(mpsz["PLN_AREA_N"])
```

# **Working with Projection**

```{r}
st_crs(mpsz)
mpsz3414 <- st_set_crs(mpsz, 3414)
st_crs(mpsz3414)
preschool3414 <- st_transform(preschool, 
                              crs = 3414)
```

# **Importing and Converting An Aspatial Data**

```{r}
listings <- read_csv("D:/KathyChiu77/ISSS624/Hands-on Ex 1/Data/aspatial/listings.csv")
list(listings) 
```

```{r}
listings_sf <- st_as_sf(listings, 
                       coords = c("longitude", "latitude"),
                       crs=4326) %>%
  st_transform(crs = 3414)
glimpse(listings_sf)
```

# **Geoprocessing with sf package**

```{r}
buffer_cycling <- st_buffer(cyclingpath, 
                               dist=5, nQuadSegs = 30)
buffer_cycling$AREA <- st_area(buffer_cycling)
sum(buffer_cycling$AREA)
mpsz3414$`PreSch Count`<- lengths(st_intersects(mpsz3414, preschool3414))
summary(mpsz3414$`PreSch Count`)
top_n(mpsz3414, 1, `PreSch Count`)
mpsz3414$Area <- mpsz3414 %>%
  st_area()
mpsz3414 <- mpsz3414 %>%
  mutate(`PreSch Density` = `PreSch Count`/Area * 1000000)
```

# **Explorotary Data Analysis (EDA)**

```{r}
hist(mpsz3414$`PreSch Density`)
ggplot(data=mpsz3414, 
       aes(x= as.numeric(`PreSch Density`)))+
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  labs(title = "Are pre-school even distributed in Singapore?",
       subtitle= "There are many planning sub-zones with a single pre-school, on the other hand, \nthere are two planning sub-zones with at least 20 pre-schools",
      x = "Pre-school density (per km sq)",
      y = "Frequency")
ggplot(data=mpsz3414, 
       aes(y = `PreSch Count`, 
           x= as.numeric(`PreSch Density`)))+
  geom_point(color="black", 
             fill="light blue") +
  xlim(0, 40) +
  ylim(0, 40) +
  labs(title = "",
      x = "Pre-school density (per km sq)",
      y = "Pre-school count")
```
