---
title: "In-class_Ex2:Spatial Weights"
author: "QIU RUILIU"
date: "25 Nov 2023"
date-modified: "last-modified"
format: html
execute: 
  eval: true
  echo: true
  warning: false
editor: visual
---

## Getting Started

### Installing and Loading the R packages

Install and load **sf**,**tmap**, **sfdep**, **tidyverse** and **knitr** packages into R environment for preparation.

```{r}
#pacman:: is for introducing pacman without installing it into R environment
pacman::p_load(sf, sfdep, knitr, tmap, tidyverse)
```

## Data Importing

For the purpose of in-class exercise, two data sets will be used:

-   Hunan county boundary layer. This is a geospatial data set in ESRI shapefile format.

-   Hunan_2012.csv: This csv file contains selected Hunan\'s local development indicators in 2012.

```{r}
hunan <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan")
```

```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```

## Combining Both Data Frame by Using Left Join

Check and grasp the basic structure of data.

```{r}
head(hunan,5)
```

```{r}
head(hunan2012,5)
```

```{r}
hunan_GDPPC <- left_join(hunan,hunan2012) %>%
  select(1:4,7,15)
```

By using `left_join`, the selected column (1, 2, 3, 4, 7, 15) of data frame **hunan** will be retained in the new one. Notably, the column **geometry** will be auto-saved for geospatial data.

Deriving Contiguity Weights: Queen's Method

In the code chunk below, queen method is used to derive the contiguity weights.

```{r}
we_q <- hunan_GDPPC %>%
  mutate(nb = st_contiguity(geometry),
         wt = st_weights(nb,
                         style = "W"),
         .before = 1)
```

Notice that `st_weights()` provides tree arguments, they are:

-   *nb*: A neighbor list object as created by `st_neighbors()`.

-   *style*:Default "W" for row standardized weights.

-   *allow_zero*:if TRUE, assign zero as lagged value to zone without neighbors.
